---
title: "AS! Case-mix trajectories"
author: "Ryan Stanley Falck"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Data Set-Up and Management

Load packages
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(readxl,psych,ggplot2,tableone,lavaan, semPlot, sem, semTools, tidyverse, haven, gbmt, dplyr, lcmm)
```

Set working directory
```{r}
setwd("~/Desktop/Manuscripts/Ongoing Projects/AS! Case-mix trajectories")
```

Load data
```{r}
falls_plus_HRU<-read_excel("AS_Econ_Falls_HRU_merge_Feb 22 2021.xlsx")
geris<-read_excel("AS Geri_July 30 2018.xlsx")
HRUs1 <- read_excel("AS! HRU Q1-3_June 13 2019.xlsx", sheet = "HRU Q1")
HRUs2 <- read_excel("AS! HRU Q1-3_June 13 2019.xlsx", sheet = "HRU Q2")
HRUs3 <- read_excel("AS! HRU Q1-3_June 13 2019.xlsx", sheet = "HRU Q3")
```

Data management of individual data frames. Preparing for merging
```{r}
falls_plus_HRU_2<-subset(falls_plus_HRU, assessmenperiod == 1)
falls_plus_HRU_2$Stroop3_2<-as.numeric(falls_plus_HRU_2$`Stroop 3`) - as.numeric(falls_plus_HRU_2$`Stroop 2`)
falls_plus_HRU_2$DSFB<-as.numeric(falls_plus_HRU_2$`Digit F`) - as.numeric(falls_plus_HRU_2$`Digit B`)
falls_plus_HRU_3<-falls_plus_HRU_2[c(1,3,4,7:10,34:36,39,40,41,45,49:51,52:54,65:71,72,76,79,82,83)]
falls_plus_HRU_3<-rename(falls_plus_HRU_3,c("ID"="id"))

geris_2<-subset(geris, ID != 330)
```

HRUQ1 Management
```{r}

#Create dataframe
HRUs1_2 <- HRUs1[c(1,2,7)]
HRUs1_3<-rename(HRUs1_2,c("Time"="Assessmen Period"))
HRUs1_4<-subset(HRUs1_3,Time!="Lagime")
HRUs1_4$Time<-as.numeric(HRUs1_4$Time)

#Delete duplicates
HRUs1_5<-distinct(HRUs1_4)

#Resolve remaining multiple values
HRUs1_6<-HRUs1_5[-c(909,3128,3129),] #Resolves double input for ID=197 Time=1. Removes both inputs for ID=369 Time=3 (inputting manually)

#Insert missing values 
ID<-c(171,171,171,251,255,255,255,255,255,369,369)
Time<-c(2,11,12,13,7,8,9,10,11,5,3)
hruq1t<-c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,563.3191)
missing<-data.frame(ID,Time,hruq1t)
HRUs1_final<-rbind(HRUs1_6, missing)
```  

HRUQ2 Management
```{r}
#Create dataframe
HRUs2_2<-rename(HRUs2,c("Time"="Assessmen Period"))
HRUs2_3<-subset(HRUs2_2,Time!="Lagime")
HRUs2_3$Time<-as.numeric(HRUs2_3$Time)

#Delete duplicates
HRUs2_4<-distinct(HRUs2_3)

#No remaining multiple values

#Insert missing values 
ID<-c(171,171,171,251,255,255,255,255,255,369)
Time<-c(2,11,12,13,7,8,9,10,11,5)
hruq2total<-c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA)
missing_2<-data.frame(ID,Time,hruq2total)
HRUs2_final<-rbind(HRUs2_4, missing_2)
```

HRUQ3 Management
```{r}
HRUs3_2<-rename(HRUs3,c("Time"="Assessmen Period"))
HRUs3_2$Time<-as.factor(HRUs3_2$Time)

#Delete duplicates
HRUs3_3<-distinct(HRUs3_2)

#Resolve remaining multiple values
HRUs3_4<-HRUs3_3[!(HRUs3_3$ID=="369" & HRUs3_3$Time=="3"),]#Removes both inputs for ID=369 Time=3 (inputting manually)

ID<-c(171,171,171,251,255,255,255,255,255,369,369)
Time<-c(2,11,12,13,7,8,9,10,11,5,3)
hruq3total<-c(NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,314.7633)
missing_3<-data.frame(ID,Time,hruq3total)
HRUs3_final<-rbind(HRUs3_4, missing_3)
```

Merge HRU data
```{r}
HRUstotal_1<- merge(HRUs1_final,HRUs2_final)
HRUstotal_final<- merge(HRUstotal_1,HRUs3_final)
```

Create HRU Total Cost
```{r}
#HRU total cost across three categories
HRUstotal_final$HRU_totalcost<-HRUstotal_final$hruq1t + HRUstotal_final$hruq2total + HRUstotal_final$hruq3total

#Accounting for inflation
HRUstotal_final$HRU_totalcost_inflated<-HRUstotal_final$HRU_totalcost *1.1110

#Because the costs are large numbers, this will shrink the dependent variable and hopefully improve model fit
HRUstotal_final$HRU_totalcost_small <- HRUstotal_final$HRU_totalcost_inflated/1000

#As another alternative I normalized cost data in order to account for the issues of non-normality using a min-max normalization
HRUstotal_final$HRU_normalized <- (HRUstotal_final$HRU_totalcost_small - min(HRUstotal_final$HRU_totalcost_small, na.rm=TRUE)) / (max(HRUstotal_final$HRU_totalcost_small, na.rm = TRUE) - min(HRUstotal_final$HRU_totalcost_small, na.rm = TRUE))

#I also log-transformed cost data in order to account for the issues of non-normality using a min-max normalization
HRUstotal_final$HRU_log <- log(HRUstotal_final$HRU_totalcost_inflated + 1)
```


Create final data frame for all variables
```{r}
long.data<-merge(falls_plus_HRU_3, HRUstotal_final)
```

Final Data Management
```{r}
long.data2<-long.data[c(1:3,8,33,38:41)]
long.data2$Group<-as.factor(long.data2$Group)
```

Note: Two participants did not provide HRU data (ID= 330 and ID=450). All other participants were included in the model.

# 2) Fitting the primary model

This can be done using the lcmm package. We first explored how to fit the data according to the link function and number of polynomials (up to 3 orders). We examined: 1) a linear link function; 2) a spline function using 5 equidistant splines; and 3) a spline function using 3 quantiles. We do not include covariates because this causes issues with the Hessian (i.e., the estimate of the curvature of the log-likelihood surface at the maximum likelihood estimate) - this is likely due to complete separation wherein one or more parameters perfectly predict the response parameter 
```{r}
model1<-lcmm(HRU_totalcost_small~Time, random=~ Time, subject='ID', ng = 1 , link = 'linear', data = long.data2)
model2<-lcmm(HRU_totalcost_small~Time, random=~Time, subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model3<-lcmm(HRU_totalcost_small~Time, random=~Time, subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model4<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~ Time + I(Time^2), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model5<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model6<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model7<-lcmm(HRU_totalcost_small~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model8<-lcmm(HRU_totalcost_small~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model9<-lcmm(HRU_totalcost_small~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)

summarytable(model1, model2, model3, model4, model5, model6, model7, model8, model9, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```
We then compared the estimated functions by plot.
```{r}
col <- rainbow(10)
plot(model1, which="linkfunction",bty='l', ylab="HRUCost", col=col[1], lwd=2, xlab="underlying latent process")
plot(model2, which="linkfunction", add=TRUE, col=col[2], lwd=2)
plot(model3, which="linkfunction", add=TRUE, col=col[3], lwd=2)
plot(model4, which="linkfunction", add=TRUE, col=col[4], lwd=2)
plot(model5, which="linkfunction", add=TRUE, col=col[5], lwd=2)
plot(model6, which="linkfunction", add=TRUE, col=col[6], lwd=2)
plot(model7, which="linkfunction", add=TRUE, col=col[7], lwd=2)
plot(model8, which="linkfunction", add=TRUE, col=col[8], lwd=2)
plot(model9, which="linkfunction", add=TRUE, col=col[9], lwd=2)
legend(x="topleft", legend=c("model1", "model2","model3","model4","model5","model6","model7","model8","model9"), lty=1, col=col, bty="n", lwd=2)
```

The results suggest that either a 2nd or 3rd order model using 3 splines at 25%, 50% and 75% provides the best fit. Next we investigate the number of groups in each model.

First order models are described below. The best fitting models is model3d (4 classes). 
```{r}
model3a<-lcmm(HRU_totalcost_small~Time, random=~Time, subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model3b<-lcmm(HRU_totalcost_small~Time, random=~Time, mixture =~ Time, subject='ID', ng = 2 , link = '3-quant-splines', data = long.data2, B=model3)
model3c<-lcmm(HRU_totalcost_small~Time, random=~Time, mixture =~ Time, subject='ID', ng = 3 , link = '3-quant-splines', data = long.data2, B=model3)
model3d<-lcmm(HRU_totalcost_small~Time, random=~Time, mixture =~ Time, subject='ID', ng = 4 , link = '3-quant-splines', data = long.data2, B=model3)
model3e<-lcmm(HRU_totalcost_small~Time, random=~Time, mixture =~ Time, subject='ID', ng = 5 , link = '3-quant-splines', data = long.data2, B=model3)
model3f<-lcmm(HRU_totalcost_small~Time, random=~Time, mixture =~ Time, subject='ID', ng = 6 , link = '3-quant-splines', data = long.data2, B=model3)

summarytable(model3a, model3b, model3c, model3d, model3e, model3f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```


Second order models are described below. Model6c (3 classes), model6d (4 classes), model6e (5 classes), and model6f (6 classes) are the best fitting models
```{r}
model6a<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model6b<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), mixture =~ Time + I(Time^2), subject='ID', ng = 2 , link = '3-quant-splines', data = long.data2, B=model6)
model6c<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), mixture =~ Time + I(Time^2), subject='ID', ng = 3 , link = '3-quant-splines', data = long.data2, B=model6)
model6d<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), mixture =~ Time + I(Time^2), subject='ID', ng = 4 , link = '3-quant-splines', data = long.data2, B=model6)
model6e<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), mixture =~ Time + I(Time^2), subject='ID', ng = 5 , link = '3-quant-splines', data = long.data2, B=model6)
model6f<-lcmm(HRU_totalcost_small~Time + I(Time^2), random=~Time + I(Time^2), mixture =~ Time + I(Time^2), subject='ID', ng = 6 , link = '3-quant-splines', data = long.data2, B=model6)

summarytable(model6a, model6b, model6c, model6d, model6e, model6f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```


The best fitting models are then compared here
```{r}
summarytable(model3d, model6c, model6d, model6e, model6f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

Summary plots comparing models are here
```{r}
summaryplot(model3d, model6c, model6d, model6e, model6f, which = c("AIC", "BIC", "entropy", "ICL"))
```

Based on these results, it seems that each model is somewhat comparable. Thus, we plotted each of the case-mix trajectories in order to determine which one appeared to best explain the data

Model 3d: 1st order with 4 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model3d, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green", "purple"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per $1000)", legend=NULL, main="Linear Model: 4 Classes",ylim=c(0,5))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3", "Class 4"), col=c("red","navy", "green",  "purple"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```


Model 6C: 2nd order with 3 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model6c, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per $1000)", legend=NULL,  main="Quadratic Model: 3 Classes",ylim=c(0,10))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3"), col=c("red","navy", "green"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```


Model 6D: 2nd order with 4 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model6d, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green", "purple"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per $1000)",legend=NULL,main="Quadratic Model: 4 Classes",ylim=c(0,10))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3", "Class 4"), col=c("red","navy", "green", "purple"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```


Model 6E: 2nd order with 5 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model6e, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green", "purple", "gold"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per$1000)",legend=NULL,  main="Quadratic Model: 5 Classes",ylim=c(0,10))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3", "Class 4", "Class 5"), col=c("red","navy", "green", "purple", "gold"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Model 6F: 2nd order with 6 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model6f, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green", "purple", "gold", "magenta"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per$1000)",legend=NULL,  main="Quadratic Model: 6 Classes",ylim=c(0,10))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3", "Class 4", "Class 5", "Class 6"), col=c("red","navy", "green", "purple", "gold", "magenta"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

## 2.1) Final model

Based on these data, the final model was a linear 4 class model using 3-quant splines.

```{r}
summary(model3d)
```

```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model3d, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green", "purple"), lwd = 2, xlab="Time (Months)", ylab="HRU Total Cost (per $1000)", legend=NULL, main="Case-mix trajectories for Health Resource Utilization (2021 CAD)",ylim=c(0,5))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3", "Class 4"), col=c("red","navy", "green",  "purple"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Plot in ggplot2
```{r}
OEP_pred <- predictY(model3d, OEP_data, var.time ="Time", draws = FALSE)
Time<-OEP_pred$times
Classes<-as.data.frame(OEP_pred$pred)
OEP_data_frame<-cbind(Time, Classes)

Class1_frame<-OEP_data_frame[c(1,2)]
Class2_frame<-OEP_data_frame[c(1,3)]
Class3_frame<-OEP_data_frame[c(1,4)]
Class4_frame<-OEP_data_frame[c(1,5)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure1<-ggplot(graph_data, aes(x=Time, y=Predicted, group=Class)) + geom_smooth(aes(colour=Class)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 1 - case trajectories.tiff",figure1, width = 4, height = 4, dpi = 300)
```



Predicted posterior probabilities
```{r}
postprob(model3d)
```


### 2.2) Describe classes
Create data frame with classes
```{r}
model3d.frame<-model3d$pprob

baseline.data<-merge(model3d.frame, falls_plus_HRU_3, by="ID")

baseline.data.2 <- rename(baseline.data, "Falls_number" = "Total Falls", "Final_Weight" = "Final Weight", "Final_Height"="Final Height", "WHR"="Waist to Hip Ratio","FCI"="FCI Total",
                          "IADL"="Total IADL", "TrailsBA"="Trail Difference","TUG"="Mean TUG", "GaitSpeed"="Gait Speed", "SPPB"="NIA Total Score") 

baseline.data.2$Gender[baseline.data.2$Gender==1]<-"Female"
baseline.data.2$Gender[baseline.data.2$Gender==2]<-"Male"

baseline.data.2$BMI<-baseline.data.2$Final_Weight/((baseline.data.2$Final_Height/100)^2)


```

Describe participants
```{r}
baseline.vars<-dput(names(baseline.data.2[c(8,13,7,17,16,38,20,21,23,12,33,34,35,24,25,36,37)]))
baseline.data.2$Group<-as.factor(baseline.data.2$Group)

Table1_full<-CreateTableOne(vars=baseline.vars, data=baseline.data.2)
print(Table1_full,contDigits=2,missing=TRUE,quote=TRUE)

Table1_stratified<-CreateTableOne(vars=baseline.vars, strata="class", data=baseline.data.2)
print(Table1_stratified,contDigits=2,missing=TRUE,quote=TRUE)
```

# 3) Moderation analysis


Moderation data set-up for cognitive and physical function
```{r}
long.data3<-long.data[c(1,19,20,30,33,38)]
long.data3$HRU_totalcost_small <- long.data3$HRU_totalcost_inflated/1000

#Binary variable for MCI (1=MCI)
long.data3$MCI<-NA
long.data3$MCI[long.data3$MoCA<26] <- 1
long.data3$MCI[long.data3$MoCA>=26] <- 0

#Binary variable for SPPB (1=Physical frailty)
long.data3$Frailty<-NA
long.data3$Frailty[long.data3$`NIA Total Score`<9] <- 1
long.data3$Frailty[long.data3$`NIA Total Score`>=9] <- 0

long.data4<-long.data3[c(1,5:9)]
```

## 3.1) Moderation by group

Model development
```{r}
long.data2$Group.1<-NA
long.data2$Group.1[long.data2$Group==0]<-"UC"
long.data2$Group.1[long.data2$Group==1]<-"OEP"


#Base model
group.model.base<-lcmm(HRU_totalcost_small~Time*Group.1, random=~Time, subject='ID', ng = 1, link = '3-quant-splines', data = long.data2)

#Full model
group.model.moderation<-lcmm(HRU_totalcost_small~Time*Group.1, random=~Time, subject='ID', mixture =~ Time*Group.1, ng = 4, link = '3-quant-splines', data = long.data2, B=group.model.base)
```

Base model summary
```{r}
summary(group.model.base)
```

Full moderation model summary
```{r}
summary(group.model.moderation)
```

Graph
```{r}
OEP_mod_data<-data.frame(Time= seq(1,13, length=1000), Group.1=rep(c("OEP","UC")))
OEP_mod_pred <- predictY(group.model.moderation, OEP_mod_data, var.time ="Time", draws = FALSE)

Time<-OEP_mod_pred$times
Group <- rep(c('OEP', 'UC'), length.out = 1000)
Classes<-as.data.frame(OEP_mod_pred$pred)
OEP_mod_data_frame<-cbind(Time, Group, Classes)

Class1_frame<-OEP_mod_data_frame[c(1,2,3)]
Class2_frame<-OEP_mod_data_frame[c(1,2,4)]
Class3_frame<-OEP_mod_data_frame[c(1,2,5)]
Class4_frame<-OEP_mod_data_frame[c(1,2,6)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure2A<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=Group)) + geom_smooth(aes(colour=Class, linetype=Group)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2A - case trajectories.tiff",figure2A, width = 4, height = 4, dpi = 300)
```


## 3.2) Moderation by age

Model development
```{r}
#Summarize Age variable
summary(long.data2$Age) #25% Quartile = 77; 75% Quartile = 86; Median = 82

#Create binary age variable (i.e., > median age)
long.data2$Age_binary<-NA
long.data2$Age_binary[long.data2$Age<=82] <- "High"
long.data2$Age_binary[long.data2$Age>82]<-"Low"

#Base model
age.model.base<-lcmm(HRU_totalcost_small~Time*Age_binary, random=~Time, subject='ID', ng = 1, link = '3-quant-splines', data = long.data2)

#Full model
age.model.moderation<-lcmm(HRU_totalcost_small~Time*Age_binary, random=~Time, subject='ID', mixture =~ Time*Age_binary, ng = 4, link = '3-quant-splines', data = long.data2, B=age.model.base)
```

Base model summary
```{r}
summary(age.model.base)
```

Full moderation model summary 
```{r}
summary(age.model.moderation)
```

Graph
```{r}
Age_mod_data<-data.frame(Time= seq(1,13, length=1000), Age_binary=rep(c("High","Low")))
Age_mod_pred <- predictY(age.model.moderation, Age_mod_data, var.time ="Time", draws = FALSE)

Time<-Age_mod_pred$times
Age_binary <- rep(c('High', 'Low'), length.out = 1000)
Classes<-as.data.frame(Age_mod_pred$pred)
Age_mod_data_frame<-cbind(Time, Age_binary, Classes)

Class1_frame<-Age_mod_data_frame[c(1,2,3)]
Class2_frame<-Age_mod_data_frame[c(1,2,4)]
Class3_frame<-Age_mod_data_frame[c(1,2,5)]
Class4_frame<-Age_mod_data_frame[c(1,2,6)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure2B<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=Age_binary)) + geom_smooth(aes(colour=Class, linetype=Age_binary)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2B - case trajectories.tiff",figure2B, width = 4, height = 4, dpi = 300)
```


## 3.3) Moderation by Sex

Model development
```{r}
long.data2$Sex<-NA
long.data2$Sex[long.data2$Gender==2]<- "Male"
long.data2$Sex[long.data2$Gender==1]<- "Female"


#Base model
sex.model.base<-lcmm(HRU_totalcost_small~Time*Sex, random=~Time, subject='ID', ng = 1, link = '3-quant-splines', data = long.data2)

#Full model
sex.model.moderation<-lcmm(HRU_totalcost_small~Time*Sex, random=~Time, subject='ID', mixture =~ Time*Sex, ng = 4, link = '3-quant-splines', data = long.data2, B=age.model.base)
```

Base model summary
```{r}
summary(sex.model.base)
```

Full moderation model summary
```{r}
summary(sex.model.moderation)
```

Graph
```{r}
Sex_mod_data<-data.frame(Time= seq(1,13, length=1000), Sex=rep(c("Male","Female")))
Sex_mod_pred <- predictY(sex.model.moderation, Sex_mod_data, var.time ="Time", draws = FALSE)

Time<-Age_mod_pred$times
Sex <- rep(c('Male', 'Female'), length.out = 1000)
Classes<-as.data.frame(Sex_mod_pred$pred)
Sex_mod_data_frame<-cbind(Time, Sex, Classes)

Class1_frame<-Sex_mod_data_frame[c(1,2,3)]
Class2_frame<-Sex_mod_data_frame[c(1,2,4)]
Class3_frame<-Sex_mod_data_frame[c(1,2,5)]
Class4_frame<-Sex_mod_data_frame[c(1,2,6)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure2C<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=Sex)) + geom_smooth(aes(colour=Class, linetype=Sex)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2C - case trajectories.tiff",figure2C, width = 4, height = 4, dpi = 300)
```


## 3.4) Moderation by cognitive function

Model development
```{r}
long.data4$MCI.2<-NA
long.data4$MCI.2[long.data4$MCI==1]<-"MCI"
long.data4$MCI.2[long.data4$MCI==0]<-"No MCI"

#Base model
MCI.model.base<-lcmm(HRU_totalcost_small~Time*MCI.2, random=~Time, subject='ID', ng = 1, link = '3-quant-splines', data = long.data4)

#Full model
MCI.model.moderation<-lcmm(HRU_totalcost_small~Time*MCI.2, random=~Time, subject='ID', mixture =~ Time*MCI.2, ng = 4, link = '3-quant-splines', data = long.data4, B=age.model.base)
```

Base model summary
```{r}
summary(MCI.model.base)
```

Full moderation model summary
```{r}
summary(MCI.model.moderation)
```

Graph
```{r}
Cog_mod_data<-data.frame(Time= seq(1,13, length=1000), MCI.2=rep(c("MCI","No MCI")))
Cog_mod_pred <- predictY(MCI.model.moderation, Cog_mod_data, var.time ="Time", draws = FALSE)

Time<-Cog_mod_pred$times
MCI.2 <- rep(c('MCI', 'No MCI'), length.out = 1000)
Classes<-as.data.frame(Cog_mod_pred$pred)
Cog_mod_data_frame<-cbind(Time, MCI.2, Classes)

Class1_frame<-Cog_mod_data_frame[c(1,2,3)]
Class2_frame<-Cog_mod_data_frame[c(1,2,4)]
Class3_frame<-Cog_mod_data_frame[c(1,2,5)]
Class4_frame<-Cog_mod_data_frame[c(1,2,6)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure2d<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=MCI.2)) + geom_smooth(aes(colour=Class, linetype=MCI.2)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2D - case trajectories.tiff",figure2d, width = 4, height = 4, dpi = 300)
```
## 3.4) Moderation by physical function

Model development
```{r}
long.data4$Frailty.2<-NA
long.data4$Frailty.2[long.data4$Frailty==1]<-"Frail"
long.data4$Frailty.2[long.data4$Frailty==0]<-"Non-Frail"

#Base model
Frailty.model.base<-lcmm(HRU_totalcost_small~Time*Frailty.2, random=~Time, subject='ID', ng = 1, link = '3-quant-splines', data = long.data4)

#Full model
Frailty.model.moderation<-lcmm(HRU_totalcost_small~Time*Frailty.2, random=~Time, subject='ID', mixture =~ Time*Frailty.2, ng = 4, link = '3-quant-splines', data = long.data4, B=age.model.base)
```

Base model summary
```{r}
summary(Frailty.model.base)
```

Full moderation model summary 
```{r}
summary(Frailty.model.moderation)
```

Graph
```{r}
Frailty_mod_data<-data.frame(Time= seq(1,13, length=1000), Frailty.2=rep(c("Frail","Non-Frail")))
Frailty_mod_pred <- predictY(Frailty.model.moderation, Frailty_mod_data, var.time ="Time", draws = FALSE)

Time<-Frailty_mod_pred$times
Frailty.2 <- rep(c('Frail', 'Non-Frail'), length.out = 1000)
Classes<-as.data.frame(Frailty_mod_pred$pred)
Frailty_mod_data_frame<-cbind(Time, Frailty.2, Classes)

Class1_frame<-Frailty_mod_data_frame[c(1,2,3)]
Class2_frame<-Frailty_mod_data_frame[c(1,2,4)]
Class3_frame<-Frailty_mod_data_frame[c(1,2,5)]
Class4_frame<-Frailty_mod_data_frame[c(1,2,6)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"
Class3_frame$Class<- "Class 3"
Class4_frame$Class<- "Class 4"
Class1_frame <- rename(Class1_frame, "Predicted" = "Ypred_class1")
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class3_frame <- rename(Class3_frame, "Predicted" = "Ypred_class3")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class1_frame, Class2_frame, Class3_frame, Class4_frame)

figure2e<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=Frailty.2)) + geom_smooth(aes(colour=Class, linetype=Frailty.2)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2E - case trajectories.tiff",figure2e, width = 4, height = 4, dpi = 300)
```

# 4) Sensitivity Analyses

As a sensitivity analysis we first determined model fit with the normalized or log-transformed HRU data

## 4.1) Model fit comparisons

### 4.1.1) Normalized data
```{r}
model10<-lcmm(HRU_normalized~Time, random=~ Time, subject='ID', ng = 1 , link = 'linear', data = long.data2)
model11<-lcmm(HRU_normalized~Time, random=~Time, subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model12<-lcmm(HRU_normalized~Time, random=~Time, subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model13<-lcmm(HRU_normalized~Time + I(Time^2), random=~ Time + I(Time^2), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model14<-lcmm(HRU_normalized~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model15<-lcmm(HRU_normalized~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model16<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model17<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model18<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)

summarytable(model10, model11, model12, model13, model14, model15, model16, model17, model18, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

The best fitting models were cubic linear (model16), cubic 3-equi-spline (model17), and cubic 3-quanitative splines (model18)

Cubic spline linear models are described belowl. None of these models fits very well.
```{r}
model16a<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model16b<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3),  mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 2 , link = 'linear', data = long.data2, B=model16a)
model16c<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 3 , link = 'linear', data = long.data2, B=model16a)
model16d<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 4 , link = 'linear', data = long.data2, B=model16a)
model16e<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 5 , link = 'linear', data = long.data2, B=model16a)
model16f<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 6 , link = 'linear', data = long.data2, B=model16a)

summarytable(model16a, model16b, model16c, model16d, model16e, model16f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

Cubic equi spline models are described below. The only reasonably good fitting models are the 2 class (model 17b) and 3 class (model 17c) models 
```{r}
model17a<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model17b<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3),  mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 2 , link = '3-equi-splines', data = long.data2, B=model17a)
model17c<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 3 , link = '3-equi-splines', data = long.data2, B=model17a)
model17d<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 4 , link = '3-equi-splines', data = long.data2, B=model17a)
model17e<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 5 , link = '3-equi-splines', data = long.data2, B=model17a)
model17f<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 6 , link = '3-equi-splines', data = long.data2, B=model17a)

summarytable(model17a, model17b, model17c, model17d, model17e, model17f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

Cubic quant-spline models are described below. No model fits the data very well
```{r}
model18a<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model18b<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3),  mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 2 , link = '3-quant-splines', data = long.data2, B=model18a)
model18c<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 3 , link = '3-quant-splines', data = long.data2, B=model18a)
model18d<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 4 , link = '3-quant-splines', data = long.data2, B=model18a)
model18e<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 5 , link = '3-quant-splines', data = long.data2, B=model18a)
model18f<-lcmm(HRU_normalized~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), mixture =~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 6 , link = '3-quant-splines', data = long.data2, B=model18a)

summarytable(model18a, model18b, model18c, model18d, model18e, model18f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

We then plotted the only two models which seemed to fit the data reasonably well

Cubic, equi-spline model with 2 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model17b, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy"), lwd = 2, xlab="Time (Months)", ylab="Normalized HRU Total Cost", legend=NULL, main="Cubic Model, Equispline: 2 Classes",ylim=c(0,.6))
legend(x="topleft",legend=c("Class 1","Class 2"), col=c("red","navy"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Cubic, equi-spline model with 3 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model17c, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green"), lwd = 2, xlab="Time (Months)", ylab="Normalized HRU Total Cost", legend=NULL, main="Cubic Model, Equi-Spline: 3 Classes",ylim=c(0,.6))
legend(x="topleft",legend=c("Class 1","Class 2", "Class 3"), col=c("red","navy", "green"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Neither of these models fit well

### 4.1.2) Log-transformed data
```{r}
model19<-lcmm(HRU_log~Time, random=~ Time, subject='ID', ng = 1 , link = 'linear', data = long.data2)
model20<-lcmm(HRU_log~Time, random=~Time, subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model21<-lcmm(HRU_log~Time, random=~Time, subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model22<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model23<-lcmm(HRU_log~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model24<-lcmm(HRU_log~Time + I(Time^2), random=~Time + I(Time^2), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model25<-lcmm(HRU_log~Time + I(Time^2) + I(Time^3), random=~ Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = 'linear', data = long.data2)
model26<-lcmm(HRU_log~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model27<-lcmm(HRU_log~Time + I(Time^2) + I(Time^3), random=~Time + I(Time^2) + I(Time^3), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)

summarytable(model19, model20, model21, model22, model23, model24, model25, model26, model27, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

The best fitting models were quadratic 3-quant spline (model 24), 1st order 3-quant spline (model 21), and quadratic 3-equi-spline (model 23)

Quadratic 3-quant spline models are described below. The best fitting models were 2 class (model 24b) and 3 class models (model 24c).
```{r}
model24a<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2), subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model24b<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 2 , link = '3-quant-splines', data = long.data2, B=model24a)
model24c<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 3 , link = '3-quant-splines', data = long.data2, B=model24a)
model24d<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 4 , link = '3-quant-splines', data = long.data2, B=model24a)
model24e<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 5 , link = '3-quant-splines', data = long.data2, B=model24a)
model24f<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 6 , link = '3-quant-splines', data = long.data2, B=model24a)

summarytable(model24a, model24b, model24c, model24d, model24e, model24f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```


1st order 3-quant spline models are described below. The best fitting models was the 2 class model (model 21b)
```{r}
model21a<-lcmm(HRU_log~Time, random=~ Time, subject='ID', ng = 1 , link = '3-quant-splines', data = long.data2)
model21b<-lcmm(HRU_log~Time, random=~ Time, subject='ID', mixture =~ Time, ng = 2 , link = '3-quant-splines', data = long.data2, B=model21a)
model21c<-lcmm(HRU_log~Time, random=~ Time, subject='ID', mixture =~ Time, ng = 3 , link = '3-quant-splines', data = long.data2, B=model21a)
model21d<-lcmm(HRU_log~Time, random=~ Time, subject='ID', mixture =~ Time, ng = 4 , link = '3-quant-splines', data = long.data2, B=model21a)
model21e<-lcmm(HRU_log~Time, random=~ Time, subject='ID', mixture =~ Time, ng = 5 , link = '3-quant-splines', data = long.data2, B=model21a)
model21f<-lcmm(HRU_log~Time, random=~ Time, subject='ID', mixture =~ Time, ng = 6 , link = '3-quant-splines', data = long.data2, B=model21a)

summarytable(model21a, model21b, model21c, model21d, model21e, model21f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```


Quadratic 3-equi spline models are described below. The best fitting model was the 2 class model (model 23b)
```{r}
model23a<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2), subject='ID', ng = 1 , link = '3-equi-splines', data = long.data2)
model23b<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 2 , link = '3-equi-splines', data = long.data2, B=model23a)
model23c<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 3 , link = '3-equi-splines', data = long.data2, B=model23a)
model23d<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 4 , link = '3-equi-splines', data = long.data2, B=model23a)
model23e<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 5 , link = '3-equi-splines', data = long.data2, B=model23a)
model23f<-lcmm(HRU_log~Time + I(Time^2), random=~ Time + I(Time^2),  mixture =~ Time + I(Time^2), subject='ID', ng = 6 , link = '3-equi-splines', data = long.data2, B=model23a)

summarytable(model23a, model23b, model23c, model23d, model23e, model23f, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```

Here is a comparison of the best fitting models.
```{r}
summarytable(model24b, model24c, model21b, model23b, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))
```


We then plotted the the trajectories of each model

Quadratic 3-quant spline with 2 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model24b, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy"), lwd = 2, xlab="Time (Months)", ylab="Log-transformed HRU Total Cost (per $1000)", legend=NULL, main="Quadratic 3-quant spline mode: 2 Classes",ylim=c(0,5))
legend(x="topleft",legend=c("Class 1","Class 2"), col=c("red","navy"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Quadratic 3-quant spline with 3 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model24c, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy", "green"), lwd = 2, xlab="Time (Months)", ylab="Log-transformed HRU Total Cost (per $1000)", legend=NULL, main="Quadratic 3-quant spline model: 3 Classes",ylim=c(0,8))
legend(x="topleft",legend=c("Class 1","Class 2"), col=c("red","navy", "green"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

1st order 3-quant spline with 2 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model21b, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy"), lwd = 2, xlab="Time (Months)", ylab="Log-transformed HRU Total Cost (per $1000)", legend=NULL, main="1st order 3-quant spline model: 2 Classes",ylim=c(0,8))
legend(x="topleft",legend=c("Class 1","Class 2"), col=c("red","navy"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

Quadratic 3-quant spline with 2 classes
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model23b, OEP_data, var.time ="Time", draws = FALSE)
plot(OEP_pred, col=c("red","navy"), lwd = 2, xlab="Time (Months)", ylab="Log-transformed HRU Total Cost (per $1000)", legend=NULL, main="Quadratic 3-quant spline model: 2 Classes",ylim=c(0,8))
legend(x="topleft",legend=c("Class 1","Class 2"), col=c("red","navy"), lty=1,lwd=2, ncol=1, bty="n", cex = 0.7)
```

### 4.1.3) Final Model

Based on these data, the most simple and easily interpretable model is model21b.

```{r}
summary(model21b)
postprob(model21b)
```

Graph of model 21b
```{r}
OEP_data<-data.frame(Time= seq(1,13, length=1000))
OEP_pred <- predictY(model21b, OEP_data, var.time ="Time", draws = FALSE)
Time<-OEP_pred$times
Classes<-as.data.frame(OEP_pred$pred)
OEP_data_frame<-cbind(Time, Classes)
OEP_data_frame$class1_untrans<-exp(OEP_data_frame$Ypred_class1)
OEP_data_frame$class2_untrans<-exp(OEP_data_frame$Ypred_class2)

Class1_frame<-OEP_data_frame[c(1,4)]
Class2_frame<-OEP_data_frame[c(1,5)]
Class1_frame$Class<- "Class 1"
Class2_frame$Class<- "Class 2"

Class1_frame<-rename(Class1_frame,c("HRU"="class1_untrans"))
Class2_frame<-rename(Class2_frame,c("HRU"="class2_untrans"))

graph_data<-rbind(Class1_frame, Class2_frame)

ggplot(graph_data, aes(x=Time, y=HRU, group=Class)) + geom_line(aes(colour=Class)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost") + ylim(0,250)
```

### 4.1.4) Class descriptions
Create data frame with classes
```{r}
model21b.frame<-model21b$pprob

baseline.data.3<-merge(model21b.frame, falls_plus_HRU_3, by="ID")

baseline.data.4 <- rename(baseline.data.3, "Falls_number" = "Total Falls", "Final_Weight" = "Final Weight", "Final_Height"="Final Height", "WHR"="Waist to Hip Ratio","FCI"="FCI Total",
                          "IADL"="Total IADL", "TrailsBA"="Trail Difference","TUG"="Mean TUG", "GaitSpeed"="Gait Speed", "SPPB"="NIA Total Score") 

baseline.data.4$Gender[baseline.data.4$Gender==1]<-"Female"
baseline.data.4$Gender[baseline.data.4$Gender==2]<-"Male"

baseline.data.4$BMI<-baseline.data.4$Final_Weight/((baseline.data.4$Final_Height/100)^2)
```

Describe participants
```{r}
baseline.vars.2<-dput(names(baseline.data.4[c(11,5,15,14,36,18,19,21,10,31,32,34,22,23,34,35)]))

Table1_full<-CreateTableOne(vars=baseline.vars.2, strata="class", data=baseline.data.4)
print(Table1_full,contDigits=2,missing=TRUE,quote=TRUE)
```


# 5) Figure 1 REVISED

Exclude Class 1 and Class 3 from the graph
```{r}
OEP_pred <- predictY(model3d, OEP_data, var.time ="Time", draws = FALSE)
OEP_pred_filtered <- OEP_pred$pred[, c("Ypred_class2", "Ypred_class4")]
tiff("Figure 1 REVISED.tiff", width=800, height=400)
plot(OEP_pred_filtered[,1], type="l", col="navy", lwd=2, xlab="Time (Months)", ylab="HRU Total Cost (per $1000)", 
     ylim=c(0,5)) 
lines(OEP_pred_filtered[,2], col="purple", lwd=2, lty=2)
legend(x="topleft", legend=c("Class 2", "Class 4"), col=c("navy", "purple"), lty=1, lwd=2, ncol=1, bty="n", cex=0.7)
dev.off()
```

# 6) Figure 2 REVISED

Exclude Class 1 and Class 3
```{r}
Sex_mod_data<-data.frame(Time= seq(1,13, length=1000), Sex=rep(c("Male","Female")))
Sex_mod_pred <- predictY(sex.model.moderation, Sex_mod_data, var.time ="Time", draws = FALSE)

Time<-Age_mod_pred$times
Sex <- rep(c('Male', 'Female'), length.out = 1000)
Classes<-as.data.frame(Sex_mod_pred$pred)
Sex_mod_data_frame<-cbind(Time, Sex, Classes)

Class2_frame<-Sex_mod_data_frame[c(1,2,4)]
Class4_frame<-Sex_mod_data_frame[c(1,2,6)]
Class2_frame$Class<- "Class 2"
Class4_frame$Class<- "Class 4"
Class2_frame <- rename(Class2_frame, "Predicted" = "Ypred_class2")
Class4_frame <- rename(Class4_frame, "Predicted" = "Ypred_class4")

graph_data<-rbind(Class2_frame, Class4_frame)

figure2C<-ggplot(graph_data, aes(x=Time, y=Predicted, color=Class, linetype=Sex)) + geom_smooth(aes(colour=Class, linetype=Sex)) + theme_classic() + xlab("Time (Months)") + ylab("HRU Total Cost (per $1000 CAD") + ylim(0,5)

ggsave("Figure 2 REVISED.tiff",figure2C, width = 6, height = 6, dpi = 300)
```
